---
title: "Data Preprocessing"
author: " Louise Brix Pilegaard Hansen (202006093) and Rikke Uldbæk (202007501)"
date: "2023-04-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# loading required packages
library(sp)
library(rgdal)
library(sf)
library(readxl)
library(tidyverse)

# set decimal digits
options(digits = 15)
```

# Load data from GeoJSON files

Loading the GeoJSON data in the 'data/geojson_files' folder as simple features objects:

```{r}
playgrounds_json <- st_read('../data/geojson_files/legepladser.json')

free_fitness_json <- st_read('../data/geojson_files/fitness_i_det_fri.json')

shelters_json <- st_read('../data/geojson_files/shelter.json')

climbing_json <- st_read('../data/geojson_files/træklatring.json')

forests_json <- st_read('../data/geojson_files/store_skove.json')

firepits_json <- st_read('../data/geojson_files/bålpladser.json')
```

The JSON files do not all have the same column names and contents. Therefore, we need to extract the necessary information based on the column names. This function does that, as well as adding a description, type of activity, age group and whether it is indoor or outdoor for each of the JSON frames to be used for the final map.

```{r}
create_sf <- function(df, name, act_type, description, indoor_outdoor, group){

  # save name of location based on column that contains this info
  name <- df[[name]]

  # save geometry type 
  geometry <- df$geometry

  # create activity type column
  type <- rep(act_type, length(geometry))
  
  # add constant dummy value for ShinyApp purposes
  all_type <- 'Show all'
  
  # add description column
  description <- rep(description, length(geometry))
  
  # add age group
  group <- rep(group, length(geometry))
  all_group<- 'Show all'
  
  # add indoor or outdoor column
  indoor_outdoor<- rep(indoor_outdoor, length(geometry))
  all_in_out <- 'Show all'

  # gather df with new columns
  df_new <- data.frame(name, geometry, type, all_type, description, group, all_group, indoor_outdoor, all_in_out)

  # convert to simple feature
  df_sf <- st_as_sf(df_new)

  # transform the CRS to be the same as leaflet's 
  df_sf <- df_sf %>% 
    st_transform(crs = 4326)

return(df_sf)
}
```

### Check column names

```{r}
glimpse(playgrounds_json)
glimpse(free_fitness_json)
glimpse(shelters_json)
glimpse(climbing_json)
glimpse(forests_json)
glimpse(firepits_json)
```

### Based on the column names, create new dataframes
```{r}
playgrounds <- create_sf(playgrounds_json, 'lokation', 'Playground', 'Outdoor playground', 'Outdoor', 'Children')

outdoor_fitness <- create_sf(free_fitness_json, 'navn', 'Sport and fitness','Outdoor fitness facilities', 'Outdoor', 'Adults')

shelters <- create_sf(shelters_json, 'navn', 'Nature facilities', 'Outdoor shelter', 'Outdoor', 'Both children and adults')

tree_climbing <- create_sf(climbing_json, 'navn', 'Sport and fitness', 'Tree climbing', 'Outdoor', 'Both children and adults')

forests <- create_sf(forests_json, 'navn', 'Forest', 'Forest', 'Outdoor', 'Both children and adults')

firepits <- create_sf(firepits_json, 'navn', 'Nature facilities', 'Bonfire pit in forest', 'Outdoor', 'Both children and adults')
```

## Cast all objects to 'POINTS'

When the data was loaded, we could see that most of the created sf objects were of class 'POINT', but three objects, outdoor_fitness, forests and tree_climbing are multipoints and multipolygons.

In order to get them all to be the same geometry, points, we divide the sf objects into two dataframes; one for points and one for polygons.

For some reason, the outdoor_fitness data is set as geometry type 'MULTIPOINT', even though it is only one point. We convert them into points instead:

```{r warning=FALSE}
# create dataframe from all points and multipoints objects
df_points <- rbind(playgrounds, shelters, outdoor_fitness, firepits)

# cast points+multipoints objects to 'POINTS' geometry
df_points <- st_cast(df_points, 'POINT')
```

To plot all points nicely on a leaflet map, we need to create a point to represent the polygons in our data. To do so, we calculate a geometric center of our polygons by using the st_centroid function.

```{r warning=FALSE}
# this is necessary to be able to calculate the centroids
sf_use_s2(FALSE)

# gather polygon objects
df_polygons <- rbind(tree_climbing, forests)

# calculate centroid for each object
df_polygons <- st_centroid(df_polygons)
```

All the datapoints are now represented as the 'POINT' geometry.

# Add our own excel data
```{r}
# read excel file with own Aarhus datapoints
excel_data <- read_excel("../data/collected_data.xlsx")

# convert the excel dataframe to a sf object and set CRS
excel_sf <- st_as_sf(x = excel_data, coords=c("longitude", "latitude"), crs = 4326) # use longitude and latitude columns as geometry
```

```{r}
# add all data together
df_all <- rbind(df_points, df_polygons, excel_sf)
```

To plot the points in leaflet, we need to extract the longitude and latitude from the geometries.

```{r}
# extract coordinates from geometries column
coords <- st_coordinates(df_all)

# add coordinates and rename columns to longitude and latitude
df_all <- df_all %>% 
  cbind(coords) %>% 
  rename(longitude = X,
         latitude = Y)
```


# Filter data by Postdistrikter

The current datapoints cover all of Aarhus Kommune. We are only interested in activities in the more central parts of Aarhus, so we need to do some filtering. 

```{r warning=FALSE}
# read GeoJSON file containing polygons of postal districts
postdistrikter <- st_read('../data/geojson_files/postdistrikter.json')

# transform CRS to match the other data
postdistrikter <- postdistrikter %>% 
  st_transform(crs = 4326)

# save the zipcodes we want to keep. These are all zipcodes of places close to Aarhus center. They are Aarhus C, Aarhus N, Aarhus V, Brabrand, Åbyhøj, Risskov, Egå, Viby J and Højbjerg.
zipcodes <- c(8000, 8200, 8210, 8220, 8230, 8240, 8250, 8260, 8270)

# filter by saved zipcodes
postdistrikter <- postdistrikter %>% 
  filter(postnummer %in% zipcodes)

# find all points that intersects the chosen postal districs
df_all_bounded <- st_intersection(df_all, postdistrikter)
```


```{r}
# remove unnecessary columns
df_all_bounded <- subset(df_all_bounded, select = -c(nr_dist, postdistrikt, postnummer, mi_style, mi_prinx, geometry))

# save to csv to be imported in ShinyApp script
write_csv(df_all_bounded, '../data/DiscoverAarhusData.csv')
```

